### 2014-05-29 ApTranscriptome_TR.Rmd
Added gene functional annotation using DAVID throught the RDAVIDWebService Bioconductor package

# Load libraries
library(R.utils)
library(ggplot2)
library(knitr)
library(knitcitations)
library(pander)
library(stringr)
library(data.table)
library(RCurl)
library(plyr)
library(MASS)
library(xtable)
library(methods) # required by `knitcitations` when file is `knit` file from command-line

# Load Bioconductor libraries
#source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
library(topGO) 
library(Rgraphviz)

@@

>>>>

# Load libraries
library(R.utils)
library(ggplot2)
library(knitr)
library(knitcitations)
library(pander)
library(stringr)
library(data.table)
library(RCurl)
library(plyr)
library(MASS)
library(xtable)
library(methods) # required by `knitcitations` when file is `knit` file from command-line
library(reshape2)

# Load Bioconductor libraries
#source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
#biocLite("RDAVIDWebService")
library(topGO) 
library(Rgraphviz)
library(RDAVIDWebService)

@@

## DAVID Gene Functional Annotation

As an alternative to topGO, in this section I use the [DAVID Functional Annotation web resource](http://david.abcc.ncifcrf.gov/) throught the Bioconductor [RDAVIDWebService](http://www.bioconductor.org/packages/release/bioc/html/RDAVIDWebService.html) package. Installing this packages is troublesome at best and my notes on how to do so are here ....

BLAST to *Apis mellifera* and *Drosophila* RefSeq Protein database

~~~
blastx
...
~~~

First, I registered as a user on the [DAVID webservice](http://david.abcc.ncifcrf.gov/content.jsp?file=DAVID_WebService.html).

I then connect and upload my complete transcriptome BLAST table as the gene background list.

Documentation for RDAVIDWebService is available [here](http://www.bdmg.com.ar/?page_id=695)

```{r DAVID_background}
# set up connection
david.connect <- DAVIDWebService$new(email = "johnsg@uvm.edu")

# load gene list
ApTranscriptome.Amel.Dros.refseq.prot <- read.table("results/trinity-full/blast_amel_dros_refseq_protein/ApTranscriptome_parallel_blastx_refseq_protein_amellifera_drosophila1.txt", sep = "\t", header = FALSE)
ApTranscriptome.Amel.Dros.refseq.prot <- data.table(ApTranscriptome.Amel.Dros.refseq.prot)
setnames(ApTranscriptome.Amel.Dros.refseq.prot, c("Transcript", "gi", "percID", "alnLength", "mismatchCount", "gapOpenCount", "queryStart", "queryEnd", "subjectStart", "subjectEnd", "eVal", "bitScore"))
setkey(ApTranscriptome.Amel.Dros.refseq.prot, Transcript)
str(ApTranscriptome.Amel.Dros.refseq.prot)

# extract RefSeq protein ID list
split.refseq.df <- colsplit(string = ApTranscriptome.Amel.Dros.refseq.prot$gi, pattern="\\|", names = c("gi", "ID", "ref", "XP"))
# remove version information for each gene accession
refseq.col <- strsplit(split.refseq.df$XP, split = "\\.")
refseq.id <- unlist(refseq.col)[2*1:length(refseq.col)-1]
str(refseq.id)

# add background list to DAVID
david.list <- addList(david.connect, refseq.id,
                      idType = "REFSEQ_PROTEIN",
                      listName = "ApTranscriptome_Amellifera_Drosophila_refseq_protein",
                      listType = "Background")
```

Next, I need to add candidate gene lists for each response type. Extract these from the full BLAST table.

```{r DAVID_geneList}
A22.high.geneList <- ApTranscriptome.Amel.Dros.refseq.prot[A22.high]
# most of these were were not BLASTed. extract those that did
A22.high.geneList.blast <- A22.high.geneList[which(!is.na(A22.high.geneList$gi)), ]
# extract RefSeq protein ID
A22.high.geneList.blast.XP <- colsplit(string = A22.high.geneList.blast$gi, pattern="\\|", names = c("gi", "ID", "ref", "XP"))
# remove version information for each gene accession
A22.high.geneList.refseq.col <- strsplit(A22.high.geneList.blast.XP$XP, split = "\\.")
A22.high.geneList.refseq.id <- unlist(A22.high.geneList.refseq.col)[2*1:length(A22.high.geneList.refseq.col)-1]
tail(A22.high.geneList.refseq.id)

# add genelist to DAVID
david.list <- addList(david.connect, A22.high.geneList.refseq.id,
                      idType = "REFSEQ_PROTEIN",
                      listName = "ApTranscriptome_A22_high",
                      listType = "Gene")

# set annotation category and background gene list
setAnnotationCategories(david.connect, "GOTERM_BP_ALL")
getBackgroundListNames(david.connect)
setCurrentBackgroundPosition(david.connect, 1)
getCurrentBackgroundListPosition(david.connect)

# gbtain Functional Annotation Cluster
A22.high.termCluster <- getClusterReport(david.connect, type = "Term")
head(summary(A22.high.termCluster))
members(A22.high.termCluster)
plot2D(A22.high.termCluster, 1)
```
